Simulating the Linux page cache in distributed applications with the WRENCH library

### 2020-12-09 paper_io_sim.tex

            Many simulation frameworks and applications have been developed 
            in recent years to support studies in grids, cloud, HPC, peer-to-peer 
            networks, volunteer computers.                     
            These simulators are built with models of resources such as CPU, network, 
            storage with different approaches and levels of details in different simulators. 
            The detail level of the model causes a conflict between accuracy and scalability, 
            and these simulators usually trade of one for the other 
            \cite{casanova2014simgrid}. 
            For example, some simulators are built with packet-level network models 
            such as \textit{CloudSim} and \textit{iCanCloud}, or storage model that 
            simulates the operation of storage hardware in \textit{DiskSim} 
            \cite{casanova2014simgrid}. 
            Fine-grained models allow simulators to achieve high accuracy, 
            but at the same time lead to scalability issues due to long simulation time 
            when simulated platforms and applications grow. 
            SimGrid is an open source simulation platform that has been developed 
            not only to balance the trade-off between accuracy and scalability, 
            but also to achieve versatility, which allows the use of the framework 
            to simulate applications in multiple domains \cite{casanova2014simgrid}.

            Despite the fact that many sophisticated models for resources have been 
            integrated in simulators, and the page cache shows apparent impacts on I/O
            performance, the simulation model of page cache is still missing in most 
            of simulators.
            For example, SimGrid only simulates I/O operations with storage bandwidths 
            and capacity. 
            SIMCAN is a simulation framework that models page cache by storing data 
            accessed on disk in a block cache \cite{nunez2012simcan}. 
            Page cache is also modeled in iCanCloud simulator with a component in charge 
            of managing memory accesses and cached data \cite{nunez2012icancloud}, 
            but one of the issues is scalability due to its sophisticated model.
            Ultimately, non of these simulators provide any cache write strategy, 
            cache eviction mechanism with dirty/clean data and the page cache LRU lists. 
            Cache replacement policies is applied in \cite{xu2018saving} to simulate 
            in-memory caching, but for the study in energy consumption of multi-tier 
            heterogenous network.

            In this study, we chose SimGrid to implement our model of I/O with page cache 
            due to the balancing between accuracy and scalibility, the versatility allowing it 
            to be adopted in different domains. 
            In addition, it is an actively maintained open-scource framework with 
            modular design, which enables the framework to be extended with new models. 
            The main issue users have to face when using SimGrid is everything has to be 
            implemented from scratch due to its low-level API. 
            To ease this challenge, WRENCH, another open-source library, is built on top 
            of SimGrid to abstract the low-level logic and expose a higher abstraction 
            level API \cite{casanova2018wrench}. 
            We integrate our model in WRENCH and use the SimGrid API to interact with 
            simulated storage resource.
            This approach allows us to leverage existing features in WRENCH and SimGrid.
            Firstly, we can take advantage of the process control, synchronization of 
            resource consumption by concurrent processes provided by the simulation core 
            in SimGrid. 
            Secondly, we can reuse the existing SimGrid storage model with capacity 
            and bandwidth properties, which are similar to the characteristics of memory.
            And finally, WRENCH provides a high level abstraction, which help us drop 
            the burden of implementation.

>>>>

            Many simulation frameworks exist to describe the behaviours of applications
            running on the various infrastructures available to users, such as personal computers,
            clouds and clusters.
            These simulators are built with resource models such CPU, network and 
            storage, using various different approaches to describe the hardware.
            For example, there is \textit{CloudSim} and \textit{iCanCloud} for packet-level network models 
            and \textit{DiskSim} the simulates storage hardware functioning.

            Simulators must choose between scalability and accuracy when describing the
            hardware.\cite{casanova2014simgrid}. 
            Fine-grained models allow simulators to achieve high accuracy, 
            but lead to scalability issues due to long simulation time 
            when simulated platforms and applications grow. 
            SimGrid is an open source simulation platform that has been developed 
            not only to balance the trade-off between accuracy and scalability, 
            but also to achieve versatility, allowing the use of the framework 
            to simulate applications in multiple domains. WRENCH, a workflow management
            system simulator based on the SimGrid framework, provides a high-level API
            for simulating workflows, thus facilitating the integration of new models.
            \valerie{say more about wrench}

            Despite the fact that many sophisticated models for resources have been 
            integrated in simulators, and the page cache shows apparent impacts on I/O
            performance, the simulation model of page cache is still missing in most 
            of simulators.
            For example, SimGrid only simulates I/O operations with storage bandwidths 
            and capacity. 
            SIMCAN is a simulation framework that models page cache by storing data 
            accessed on disk in a block cache \cite{nunez2012simcan}. 
            Page cache is also modeled in the iCanCloud simulator through a component which
            manages memory accesses and cached data \cite{nunez2012icancloud}. However,
            the scalability of the iCanCloud simulator is limited by its sophisticated model.
            Ultimately, none of these simulators provide any writeback cache simulator nor
            cache eviction policys through LRU lists.
            Although cache replacement policies are applied in \cite{xu2018saving} to simulate 
            in-memory caching, their simulator is specific to energy consumption of multi-tier 
            heterogenous networks.

            In this study, we use WRENCH 
            to implement our page cache model. We chose WRENCH due to accurate yet scalable implementation, resulting
            from its SimGrid backend, as well as its high-level API.
            We have integrated our model in WRENCH and use the SimGrid API to interact with 
            simulated storage resources.
            This approach allows us to leverage existing features in WRENCH and SimGrid such as process control and synchronization of 
            resource consumption.
            We also leverage the existing SimGrid storage model with capacity 
            and bandwidth properties, which are similar to the characteristics of memory.
